Care Phenotypes: A Novel Approach to Understanding Healthcare Data Collection Patterns

### 2025-03-29 docs/manuscript/manuscript.tex
V6.

>>>>

\subsection{Fairness and Bias Implementation}

We implemented a comprehensive framework for evaluating and mitigating fairness and bias in healthcare algorithms:

\subsubsection{Fairness Metrics}

Our fairness evaluation framework includes:
\begin{itemize}
    \item \textbf{Demographic Fairness}: Metrics for evaluating demographic parity across phenotypes
    \item \textbf{Clinical Fairness}: Analysis of clinical factor distributions and correlations
    \item \textbf{Fairness Visualization}: Interactive visualizations of fairness metrics and disparities
\end{itemize}

\subsubsection{Bias Detection}

The bias detection component features:
\begin{itemize}
    \item \textbf{Bias Detection Algorithms}: Methods for identifying systematic biases in care patterns
    \item \textbf{Bias Validation}: Statistical validation of detected biases
    \item \textbf{Bias Visualization}: Clear visualizations of bias patterns and their impact
\end{itemize}

\subsubsection{Bias Mitigation}

Our bias mitigation implementation includes:
\begin{itemize}
    \item \textbf{Mitigation Strategies}: Multiple approaches including reweighting, threshold adjustment, and calibration
    \item \textbf{Mitigation Validation}: Comprehensive validation of mitigation effectiveness
    \item \textbf{Mitigation Visualization}: Comparison of pre- and post-mitigation fairness metrics
\end{itemize}

\subsection{Fairness and Bias Results}

\begin{itemize}
    \item \textbf{Fairness Evaluation}:
    \begin{itemize}
        \item Reliable detection of demographic and clinical disparities
        \item Strong statistical validation of fairness metrics
        \item Clear visualization of fairness patterns across phenotypes
    \end{itemize}

    \item \textbf{Bias Detection}:
    \begin{itemize}
        \item Effective identification of systematic biases in care patterns
        \item Robust validation of bias significance
        \item Intuitive visualization of bias patterns and their impact
    \end{itemize}

    \item \textbf{Bias Mitigation}:
    \begin{itemize}
        \item Successful reduction of disparities through multiple strategies
        \item Validation of mitigation effectiveness
        \item Clear visualization of mitigation impact on fairness metrics
    \end{itemize}
\end{itemize}
